Stateless Model Checking with Data-Race Preemption Points

### 2015-10-17 pldi-16/di.tex
write about within_function and finish discussion/future section

\subsection{Suppressing malloc-related false positives}

@@

>>>>

\subsection{Suppressing ``malloc-recycle'' false positives}

@@

{\bf Restricting PPs with stack trace predicates.}
Most MCs preempt indiscriminately on any sync API call, regardless of the call-site.
However, when testing a particular module in a large codebase,
the user is likely uninterested in PPs arising from other modules.
\landslide~provides the {\tt within\_function} configuration command for a user to identify which call-sites matter most.
Before inserting a PP, \landslide~requires at least one argument to {\tt within\_function} to appear in the current thread's stack trace.
The {\tt without\_function} directive works similarly, but as a blacklist.
Multiple invocations can be used; later ones take precedence.
%\cite{landslide} provides further detail on this feature.

### 2015-10-17 pldi-16/evaluation.tex
write about within_function and finish discussion/future section

Although \quicksand~presents Iterative Deepening and data-race PPs as interconnected techniques, they each could theoretically be employed alone in other model checkers.
For example, a single-state-space tool could use data-race candidates during immediately subsequent interleavings, essentially changing the state space on the fly.
Likewise, a message-passing-only tool could employ Iterative Deepening despite data races being absent from its concurrency model.
Hence, though many of our experiments compare \quicksand~to the state-of-the-art as a whole,
we also sought to evaluate each technique individually.
Our evaluation answers the following questions:
\begin{enumerate}
	\item Does Iterative Deepening find bugs faster than SSS-MC in subset state spaces, even without data-race PPs?
			% Probably not... ICB is state of the art here.
			% In large tests, can Iterative Deepening provide partial verification by completing smaller state spaces
	\item Is combining data-race analysis with MC productive?
		\begin{enumerate}
			\item Do data-race PPs expose new bugs that couldn't be found with SSS-MC's fixed-PP-set approach?
				% Elaborate later:
				% Among those, how many were missed in a {\em completed} execution of the otherwise ``maximal'' state space?
			\item Do we avoid false positives compared to a single-execution data race analysis?
				% Explain later as:
				% How many data-race candidates were verified as benign
				% But to be fair, you have to count how many DRs are reported as "couldn't test these, check yourself" at the end.
				% Also Include:
				% How many false positives does the free-re-malloc technique suppress?
				% TODO: If you have time, re-run all of the dr-only bug tests, with DR_FALSE_NEG enabled, and see how much fewer bugs get found (how many bugs get pushed past the time limit?)
			% TODO: This one's optional. You can give up on it.
			\item Do we find data-race bugs that would be false-negatives during a single-execution analysis?%Do we avoid false negatives compared to single-pass?
				%TODO: for this experiment, set EXPLORE_BACKWARDS=0
				% TODO: And disable false-neg malloc-free technique
				% TODO: And also disable the confirmed/suspected thing
				% (where mem.c waits for reorder observed before
				% messaging the latter half of the DR to QS)
		\end{enumerate}
\end{enumerate}
\subsection{Test Suite}
% TODO: Maybe say how many lines of code total? How many lines on average per P2/pintos? (careful with pintos; lots of basecode)
Our test suite consists of \numthrlibs~``P2'' student thread libraries, from CMU's 15-410 OS class,
%across the Spring and Fall 2014 and Spring 2015 semesters;
and \numpintoses~``Pintos'' student kernels, from Berkeley's CS162 and U. Chicago's CS230 OS classes.
%
The P2 thread library comprises \texttt{thr\_create()}, \texttt{thr\_exit()}, \texttt{thr\_join()}, mutexes, condvars, semaphores, and r/w locks;
all implemented from scratch in userspace with a UNIX-like system call interface \cite{kspec,thrlib}.
%
The Pintos kernel project
involves implementing priority scheduling, \texttt{sleep()}, and user-space process management (\texttt{wait()} and \texttt{exit()})
using provided bare-bones mutex, context-switch, and virtal memory implementations
\cite{pintos}.
% P2 SLOC stats: 1807 avg; 1723 median; range 1181-4114.
% All numbers, obtained with:
% cd p2s; for i in */*; do wc -l $i/user/libthread/*.{c,h,S} $i/user/libthread/*/*.{c,h,S} ; done | grep total
% 1181 1192 1221 1230 1238 1240 1243 1261 1275 1307 1310 1318 1325 1334 1336 1345 1366
% 1388 1388 1403 1415 1416 1430 1451 1478 1498 1527 1589 1618 1635 1638 1654 1675 1676
% 1716 1719 1720 1723 1723 1727 1737 1743 1744 1751 1769 1777 1782 1789 1789 1812 1918
% 1926 1946 1994 2022 2043 2066 2077 2088 2099 2131 2164 2172 2190 2215 2227 2277 2282
% 2384 2387 2483 2486 2503 2514 2551 2597 2610 2665 4114
Though not ``real world'' programs, both projects are quite large: % maybe "complex"?
the P2s average 1807 lines of C and x86 assembly (stddev 489.5),
% Pintos SLOC stats: TODO
and the Pintoses average {\bf 9999999} % TODO

\newcommand\mxtest{\texttt{mx\_test}}
\newcommand\tej{\texttt{thr\_exit\_join}}
\newcommand\bct{\texttt{broadcast}}
\newcommand\paraguay{\texttt{paraguay}}
\newcommand\paradise{\texttt{paradise\_lost}}
\newcommand\rwl{\texttt{rwl\_test}}
We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness, \tej, a test of thread lifecycle, \bct~and \paraguay, for condvars, \paradise~for semaphores, and \rwl~for r/w locks.
For \mxtest~we enabled \landslide's mutex-testing option described in \sect{\ref{sec:landslide}}.
%, which changes the lock-set tracking to consider accesses within {\tt mutex\_lock} and {\tt mutex\_unlock} as data-race candidates.
%(Normally, they are considered protected by the very lock they implement: assuming the locks are already correct to enable productive data-race analysis on the rest of the codebase.)
\newcommand\prisema{\texttt{priority\_sema}}
\newcommand\waitsimple{\texttt{wait\_simple}}
% TODO: Add more test cases
We tested the Pintoses with 2 programs from the class test suite: \prisema, a test of the kernel scheduling algorithm, and \waitsimple, a test of process lifecycle system calls.
In total, this test suite comprises 632 unique state spaces.

@@

\subsection{Evaluating Iterative Deepening}

@@

\subsection{Finding data-race bugs missed by SSS-MC}

@@

\subsection{Avoiding false positive data-race candidates}
% Though we mechanically verify whether each data race candidate leads to a bug, each new PP can increase combinatorially..... obviously wish to avoid...

@@

One might also wonder: Why is it fair to compare the data race bugs \quicksand~finds (10 CPU-hours)
against the candidates found by a prior work's single execution (less than a minute)?
We argue this comparison is meaningful because prior work data-race tools, being not integrated with a MC,
are not intended to uncover different results under subsequent runs.
One could run a data-race tool repeatedly for 10 CPU-hours, but the advantage of stateless MC over stress testing is already well-understood.
% TODO TODO TODO get a big citation for teh above sentence.

@@

\subsection{Discussion} % and future work?

% TODO: Mention that SSSMC control never FAB in mutex_test whatsoever, while DR PPs enabled FABs.

%We tested 87 pintos kernels and found races in 45-47 of them, among which 34-47 of those bugs required data-race preemption points to expose. That means we rock.

% TODO: Future work.
% Future work: Add a way to configure even smaller subsets (eg "only mutex_locks called from site X") for cases where mx_lock and mx_unlock alone are still too big. Count the number of kernels for which this was the case.
However, larger state spaces often result from finer-grained synchronization,
which can indicate a more complicated concurrent algorithm that would require more rigorous verification than a program with a single global lock.
(analogous to a developer % changing an implementation to be finer graind and then re running the same unit test)

>>>>

Although \quicksand~presents Iterative Deepening and data-race PPs as interconnected techniques, they each could theoretically be employed alone in other model checkers.
For example, a single-state-space tool could use data-race candidates during immediately subsequent interleavings, essentially changing the state space on the fly.
Likewise, a message-passing-only tool could employ Iterative Deepening despite data races being absent from its concurrency model.
Hence, though many of our experiments compare \quicksand~to the state-of-the-art as a whole,
we also sought to evaluate each technique individually.
Our evaluation answers the following questions:
\begin{enumerate}
	\item Does Iterative Deepening find bugs faster than SSS-MC in subset state spaces, even without data-race PPs?
			% Probably not... ICB is state of the art here.
			% In large tests, can Iterative Deepening provide partial verification by completing smaller state spaces
	\item Is combining data-race analysis with MC productive?
		\begin{enumerate}
			\item Do data-race PPs expose new bugs that couldn't be found with SSS-MC's fixed-PP-set approach?
				% Elaborate later:
				% Among those, how many were missed in a {\em completed} execution of the otherwise ``maximal'' state space?
			\item Do we avoid false positives compared to a single-execution data race analysis?
				% Explain later as:
				% How many data-race candidates were verified as benign
				% But to be fair, you have to count how many DRs are reported as "couldn't test these, check yourself" at the end.
				% Also Include:
				% How many false positives does the free-re-malloc technique suppress?
				% TODO: If you have time, re-run all of the dr-only bug tests, with DR_FALSE_NEG enabled, and see how much fewer bugs get found (how many bugs get pushed past the time limit?)
			% TODO: This one's optional. You can give up on it.
			\item Do we find data-race bugs that would be false-negatives during a single-execution analysis?%Do we avoid false negatives compared to single-pass?
				%TODO: for this experiment, set EXPLORE_BACKWARDS=0
				% TODO: And disable false-neg malloc-free technique
				% TODO: And also disable the confirmed/suspected thing
				% (where mem.c waits for reorder observed before
				% messaging the latter half of the DR to QS)
		\end{enumerate}
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Test Suite}
% TODO: Maybe say how many lines of code total? How many lines on average per P2/pintos? (careful with pintos; lots of basecode)
Our test suite consists of \numthrlibs~``P2'' student thread libraries, from CMU's 15-410 OS class,
%across the Spring and Fall 2014 and Spring 2015 semesters;
and \numpintoses~``Pintos'' student kernels, from Berkeley's CS162 and U. Chicago's CS230 OS classes.
%
The P2 thread library comprises \texttt{thr\_create()}, \texttt{thr\_exit()}, \texttt{thr\_join()}, mutexes, condvars, semaphores, and r/w locks;
all implemented from scratch in userspace with a UNIX-like system call interface \cite{kspec,thrlib}.
%
The Pintos kernel project
involves implementing priority scheduling, \texttt{sleep()}, and user-space process management (\texttt{wait()} and \texttt{exit()})
using provided bare-bones mutex, context-switch, and virtal memory implementations
\cite{pintos}.
% P2 SLOC stats: 1807 avg; 1723 median; range 1181-4114.
% All numbers, obtained with:
% cd p2s; for i in */*; do wc -l $i/user/libthread/*.{c,h,S} $i/user/libthread/*/*.{c,h,S} ; done | grep total
% 1181 1192 1221 1230 1238 1240 1243 1261 1275 1307 1310 1318 1325 1334 1336 1345 1366
% 1388 1388 1403 1415 1416 1430 1451 1478 1498 1527 1589 1618 1635 1638 1654 1675 1676
% 1716 1719 1720 1723 1723 1727 1737 1743 1744 1751 1769 1777 1782 1789 1789 1812 1918
% 1926 1946 1994 2022 2043 2066 2077 2088 2099 2131 2164 2172 2190 2215 2227 2277 2282
% 2384 2387 2483 2486 2503 2514 2551 2597 2610 2665 4114
Though not ``real world'' programs, both projects are quite large: % maybe "complex"?
the P2s average 1807 lines of C and x86 assembly (stddev 489.5),
% Pintos SLOC stats: TODO
and the Pintoses average {\bf 9999999} % TODO

\newcommand\mxtest{\texttt{mx\_test}}
\newcommand\tej{\texttt{thr\_exit\_join}}
\newcommand\bct{\texttt{broadcast}}
\newcommand\paraguay{\texttt{paraguay}}
\newcommand\paradise{\texttt{paradise\_lost}}
\newcommand\rwl{\texttt{rwl\_test}}
We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness, \tej, a test of thread lifecycle, \bct~and \paraguay, for condvars, \paradise~for semaphores, and \rwl~for r/w locks.
For \mxtest, \paradise, and \paraguay, we used {\tt without\_function} to blacklist {\tt thr\_create}, {\tt thr\_exit}, and {\tt thr\_join},
and for \mxtest~we enabled \landslide's mutex-testing option
(see \sect{\ref{sec:landslide}}).
\newcommand\prisema{\texttt{priority\_sema}}
\newcommand\waitsimple{\texttt{wait\_simple}}
% TODO: Add more test cases
We tested the Pintoses with 2 programs from the class test suite: \prisema, a test of the kernel scheduling algorithm, and \waitsimple, a test of process lifecycle system calls.
For all tests, we used {\tt without\_function} to blacklist PPs on the {\tt malloc} mutex.
In total, this test suite comprises 632 unique state spaces.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Evaluating Iterative Deepening}
\label{sec:eval-id}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Finding data-race bugs missed by SSS-MC}
\label{sec:eval-drbugs}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Avoiding false positive data-race candidates}
\label{sec:eval-falsepos}
% Though we mechanically verify whether each data race candidate leads to a bug, each new PP can increase combinatorially..... obviously wish to avoid...

@@

% TODO: restructure this paragraph to account for RaceFuzzer/Portend. Be like, "Yes they explore multiple schedules, but only AFTER finding a race." And/or make this argument in prior work
One might also wonder: Why is it fair to compare the data race bugs \quicksand~finds (10 CPU-hours)
against the candidates found by a prior work's single execution (less than a minute)?
We argue this comparison is meaningful because prior work data-race tools, being not integrated with a MC,
are not intended to uncover different results under subsequent runs.
One could run a data-race tool repeatedly for 10 CPU-hours, but the advantage of stateless MC over stress testing is already well-understood.
% TODO TODO TODO get a big citation for teh above sentence.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Discussion and future work}
\label{sec:future}

{\bf Testing lock implementations.}
In \sect{\ref{sec:eval-drbugs}}, using data-race PPs compares favourably across the board to SSS-MC.
Observe in particular that in \mxtest, the control experiment found dramatically fewer bugs (just 1),
even compared to the other test cases\footnote{
% FIXME: "Aren't all lock impls assembly?" "Yes, but this one was ALL assembly."
	The one bug SSS-MC found was in a fully-assembly lock implementation. {\tt yield()}'s return value clobbered a value stored in {\tt \%eax}, which could lead to a failure after two repeated contentions. Preempting only on {\tt yield()} (in the contention loop) was sufficient.}.
%Intuitively, this is due to our control experiment being able to preempt only on the boundaries of the API which
Though for many applications of MC, assuming a correct lock implementation is sufficient,
we consider this strong evidence that any verification of low-level synchronization code must incorporate data-race PPs.

{\bf Partial verification.}
We are not the first to provide a partial verification guarantee when timing out on too-large state spaces (\sect{\ref{sec:eval-id}}).
While we guarantee safety when preempted on certain combinations of PPs,
ICB ~\cite{chess} guarantees safety under no more than a certain number of preemptions.
%according to the maximum bound reached in the time limit.
We imagine these two guarantees could be each be useful to developers in different scenarios,
and are presently working to combine the two approaches to provide both at once.
One benefit of our technique is that {\tt within\_function}-based Iterative Deepening (discussed above)
would enable expert developers to configure custom subsets of PPs they are most interested in verifying,
according to which modules of a codebase they wish to test.

% TODO: Future work.
% Future work: Add a way to configure even smaller subsets (eg "only mutex_locks called from site X") for cases where mx_lock and mx_unlock alone are still too big. Count the number of kernels for which this was the case.
{\bf Finer-grained PP subsets.}
\quicksand~was able to partially guarantee safety in {\large \bf 99999\%} of tests with large maximal state spaces.
However, in {\large \bf 1337} tests, no more than the minimal state space could be verified,
and in {\large \bf 42} tests, not even that much.
Larger state spaces often result from finer-grained locking,
which can indicate a more complicated concurrent algorithm requiring more rigorous verification than a program with a single global lock.
Hence these corner cases are important to consider for future work.
While this work used {\tt within\_function} (\sect{\ref{sec:landslide}}) {\em statically} to restrict where PPs could arise in advance of the test,
we envision future Iterative Deepening implementations could incorporate this method to {\em dynamically} subset PPs further,
making partial verification of such large tests possible.
%enabling partial verification of such large tests. % if need space

### 2015-10-17 pldi-16/intro.tex
write about within_function and finish discussion/future section

% Can I even make a claim this broad to begin with?
However, all stateless model checkers to date are bound by a fixed set of {\em preemption points}: code locations that define the granularity at which threads interleave.
% TODO: You repeat the following 1.5 sentence in the related work section; maybe you can eliminate this redundancy to save space?
For example, \textsc{CHESS} \cite{chess} preempts only on synchronization operations and library calls, which can miss lock-free shared memory races.
It provides an additional data-race analysis to report any violations of this model;
however, data-race analyses are prone to report false positives and benign races which require annotations or imprecise heuristics to reduce \cite{racerx,tsan,datacollider}.
%
On the other hand, SPIN \cite{spin}
is able to preempt threads around any shared memory access. Such fine granularity would automatically check if each data race is a real bug, but makes full state space completion intractable for all but the most rudimentary test inputs.
%
This work shows how to avoid making that tradeoff decision in advance.

>>>>

% Can I even make a claim this broad to begin with?
However, all stateless model checkers to date are bound by a fixed set of {\em preemption points}: code locations that define the granularity at which threads interleave.
% TODO: You repeat the following 1.5 sentence in the related work section; maybe you can eliminate this redundancy to save space?
For example, \textsc{CHESS} \cite{chess} preempts only on synchronization operations and library calls, which can miss lock-free shared memory races.
It provides an additional data-race analysis to report any violations of this model;
however, data-race analyses are prone to report false positives
%and benign races
which require annotations or imprecise heuristics to reduce \cite{racerx,tsan,datacollider}.
%
On the other hand, SPIN \cite{spin}
is able to preempt threads around any shared memory access. Such fine granularity would automatically check if each data race is a real bug, but makes full state space completion intractable for all but the most rudimentary test inputs.
%
This work shows how to avoid making that tradeoff decision in advance.

### 2015-10-17 pldi-16/related.tex
write about within_function and finish discussion/future section

We build upon many established model-checking techniques, dating back
% of course
to Verisoft, the original C model checker \cite{verisoft}.
%, and Eraser, the original data race detector \cite{eraser}.
%Our model checker
%\landslide~\cite{landslide} itself implements many techniques from prior work (\sect{\ref{sec:landslide}}).
%itself implements DPOR \cite{dpor},
%state space estimation \cite{estimation},
%and data-race detection \cite{eraser}.
We compare related model checkers by their treatment of shared-memory thread communication.

@@

{\bf Other techniques.} Various improvements to DPOR have been proposed, such as Dynamic Interface Reduction \cite{demeter}, Maximal Causality Reduction \cite{mcr}, and DPOR for TSO/PSO \cite{tsopso}.
These are all orthogonal to our technique.
\textsc{Parrot} \cite{parrot} combines model checking with a partially-determinizing runtime environment, but still, fewer than half the non-trivial state spaces in their evaluation could be completed, providing a strong argument for \quicksand.
Finally, Iterative Context Bounding \cite{chess} is most similar to Iterative Deepening, as both approaches provide a concrete partial verification on some subset of interleavings when full completion is intractable.
However, ICB still limits itself to a fixed set of PPs, and so far no algorithm has been proposed to dynamically add data-race PPs during a test with ICB.

>>>>

\landslide~uses many established model-checking techniques, dating back
% of course
to Verisoft, the original C model checker \cite{verisoft}.
%, and Eraser, the original data race detector \cite{eraser}.
%Our model checker
%\landslide~\cite{landslide} itself implements many techniques from prior work (\sect{\ref{sec:landslide}}).
%itself implements DPOR \cite{dpor},
%state space estimation \cite{estimation},
%and data-race detection \cite{eraser}.
We compare related model checkers by their treatment of shared-memory thread communication.

@@

{\bf Other techniques.} Various improvements to DPOR have been proposed, such as Dynamic Interface Reduction \cite{demeter}, MCR \cite{mcr}, and DPOR for TSO/PSO \cite{tsopso}.
These are all orthogonal to our technique.
\textsc{Parrot} \cite{parrot} combines MC with a partially-determinizing runtime environment, but still, fewer than half the non-trivial state spaces in their evaluation could be completed, providing a strong argument for \quicksand.
Finally, ICB \cite{chess} is most similar to Iterative Deepening, as both approaches provide a concrete partial verification on some subset of interleavings when full completion is intractable (\sect{\ref{sec:future}}).
However, ICB still limits itself to a fixed set of PPs, and so far no algorithm has been proposed to dynamically add data-race PPs during a test with ICB.
